Reducing Acceptance Marks in Emerson-Lei Automata by QBF-solving

### 2022-03-04 paper/main.tex
section 3 finished!

In the following, we define three versions of $\mcycle(\vec{e})$:
\begin{itemize}
\item $\mcycle_1(\vec{e})$ is a lightweight version, which only says
  that $\Te$ is nonempty. It does not use the information about
  automaton structure, but it comes with an interesting simplification
  of the whole formula $\Phi_{C,K}$.
\item $\mcycle_2(\vec{e})$ is an intermediate version. It says that $\Te$
  is nonempty and every transition in $\Te$ has to have a preceding
  and a succeeding transition in $\Te$, which is a necessary condition
  for being a cycle, but not a sufficient one.
\item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
  cycle. Unfortunately, it uses additional universally quantified
  variables corresponding to automata states. Transformation of
  $\Phi_{C,K}$ to prenex normal form turns the quantifiers to
  existential ones and the resulting formula thus contains quantifier
  alternation.
\end{itemize}
We write $\Phi_{i,C,K}$ when we want to emphasize that the particular
formula $\Phi_{C,K}$ contains the version $\mcycle_i(\vec{e})$.

@@

Now we present a formula optimization based on the fact that every
cycle is completely included in the transition set $\delta_S$ of some
SCC $S$. Note that we cannot reduce the number of acceptance marks in
each SCC separately as all SCCs have to share one acceptance formula.
However, for each SCC we can build a $\Phi_{C,K}$ formula separately and
then we can put all these formulas into one conjunction such that they
share the variables $\vec{i},\vec{f}$.

\todo{Honza je tu!}

For a given automaton $\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$, we first
compute all SCCs. Each SCC $S$ induces an automaton $(S,M,\Sigma,q',\delta_S,\varphi)$, where $q'$ is an arbitrary state
of $S$.

the optimization saying that So far, we have not use the fact that each cycle is A

SCC-based approach

Its first part $\xi_1(\vec{i},\vec{f})$ says that any clause does not
contain both $\Inf k$ and $\Fin k$ for the same mark $k$ as a clause
with both such terms is always unsatisfiable. Note that the subformula
$\xi_1(\vec{i},\vec{f})$ can be safely omitted, but it speeds up the QBF solving.

>>>>

In the following, we define three versions of $\mcycle(\vec{e})$:
\begin{itemize}
\item $\mcycle_1(\vec{e})$ is a lightweight version, which only says
  that $\Te$ is nonempty. It does not use the information about
  automaton structure, but it comes with an interesting simplification
  of the whole formula $\Phi_{C,K}$.
\item $\mcycle_2(\vec{e})$ is an intermediate version. It says that $\Te$
  is nonempty and every transition in $\Te$ has to have a preceding
  and a succeeding transition in $\Te$, which is a necessary condition
  for being a cycle, but not a sufficient one.
\item $\mcycle_3(\vec{e})$ is a strict version saying that $\Te$ is a
  cycle. Unfortunately, it uses additional universally quantified
  variables corresponding to automata states. Transformation of
  $\Phi_{C,K}$ to prenex normal form turns the quantifiers to
  existential ones and the resulting formula thus contains quantifier
  alternation.
\end{itemize}
We write $\Phi_{j,C,K}$ when we want to emphasize that the particular
formula $\Phi_{C,K}$ contains the version $\mcycle_j(\vec{e})$.

@@

Now we present an optimized formula construction based on the fact
that every cycle is completely included in the transition set
$\delta_S$ of some SCC $S$. Note that we cannot reduce the number of
acceptance marks in each SCC separately as all SCCs have to share the
same acceptance formula after the reduction. However, for each SCC we
can build the formula $\Phi_{C,K}$ separately and then we can put all
these formulas into one conjunction such that they share the variables
$\vec{i},\vec{f}$.

Now we describe the construction of optimized QBF formulas
$\Phi^\mathit{opt}_{j,C,K}$ formally. Given an automaton
$\mA=(Q,M,\Sigma,\delta,q_I,\varphi)$ and its SCC $S$, by $\mA^S$ we
denote the automaton corresponding to the part of $\mA$ delimited by
$S$, i.e., $\mA^S=(S,M,\Sigma,q',\delta_S,\varphi)$ where $q'$ is an
arbitrary state of $S$. Further, let $\vec{n}_S$ denote the vector of
variables $n_{t,k}$ where $t\in\delta_S$ and $k\in N_k$. By
$\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})$ we denote the formula
$\Phi_{j,C,K}$ constructed for $\mA^S$ as presented above.
Finally, we define the optimized QBF formula as
\[
  \Phi^\mathit{opt}_{j,C,K}(\vec{n}',\vec{i},\vec{f})=\bigwedge_{\textrm{SCC }S}\Phi^S_{j,C,K}(\vec{n}_S,\vec{i},\vec{f})
\]
where $\vec{n}'$ is the vector of all varibles contained in vectors
$\vec{n}_S$ for SCCs $S$.

One effect of this optimization is that the formula
$\Phi^\mathit{opt}_{j,C,K}$ contains no variables $n_{t,k}$ %and $e_t$
for transitions $t$ that do not lead between states of the same SCC.
The acceptance marks on such a transition $t$ do not affect the
acceptance of any run as $t$ appears at most once on each run. For
these transitions $t$, we can define $\rem(t)$ such that
$\mks(\rem(t))=\emptyset$.
